### 2020-07-19 article.tex
New small paragraph about EPSCISS et al.

while the driver code is part of the PETSc testsuite and available at

@@

This continuous equation is discretized with third-order Lagrange finite elements using FreeFEM. The script is available at \url{https://github.com/prj-/jolivet2020petsc}. The following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}. By default, SLEPc applies at most 5 iterations of GMRES on each column of a so-called set of active columns. Since the option \pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size of the set varies between 40 to 1. This size corresponds to the number of right-hand sides in systems that will be solved throughout LOBPCG iterations. Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM} is used, with the pseudo-block GMRES. That means that, assuming the size of the set of active columns is 40, one LOBPCG iterations implies doing at most 5 GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric} and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES iterations, using \pk{MatMult} and \pk{PCApply}. Results are gathered in~\cref{tab:lobpcg}. In the first line, the numbers of outer LOBPCG iterations are reported with the three different inner preconditioners mentioned above. \pk{PCASM} is known to yield a preconditioned operator whose condition number grows as the number of subdomains, here processes, increases. Thus, the number of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the inner solves are not converging to meaningful solutions in just 5 iterations. The multilevel preconditioners perform similarly, since the difference in the number of outer iterations is minimal (1). The second and third lines represent the number of times \pk{PCApply} and \pk{PCMatApply} are called. Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on each single active column. \pk{PCASM} and \pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}. In the fourth line, the time spent setting up the preconditioners is given. \pk{PCASM} is extremely cheap but not very robust numerically, while both multilevel preconditioners have a similar setup time. Eventually, in the last line, the time spent in the full eigensolve is given. Clearly, having both a \pk{KSP} and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to achieve reasonable performance with such an outer solver that heavily relies on blocking.

@@

@@

options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020petsc}.

@@

\section*{Acknowledgments}
The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
and B.~Smith for reviewing the successive merge requests in the PETSc
repository. This work was granted access to the GENCI-sponsored HPC resources of:
\begin{itemize}
    \item TGCC@CEA under allocation A0070607519;
    \item IDRIS@CNRS under allocation AP010611780.
\end{itemize}
Jose E.~Roman was supported by the Spanish Agencia Estatal de Investigaci\'{o}n (AEI) under project SLEPc-DA (PID2019-107379RB-I00).
% We would like to acknowledge the assistance of volunteers in putting
% together this example manuscript and supplement.

>>>>

while the driver code is part of the PETSc test suite and available at

@@

This continuous equation is discretized with third-order Lagrange finite
elements using FreeFEM. The script \texttt{blocking-slepc.edp} available at
\url{https://github.com/prj-/jolivet2020petsc} can be use for reproducibility.
The following timings and convergence histories have been obtained on $2{,}304$
processes of Ir\`ene, for solving a problem of dimension
\pgfmathprintnumber{74618461}. By default, SLEPc applies at most 5 iterations
of GMRES on each column of a so-called set of active columns. Since the option
\pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size
of the set varies between 40 to 1. This size corresponds to the number of
right-hand sides in systems that will be solved throughout LOBPCG iterations.
Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM}
is used, with the pseudo-block GMRES. That means that, assuming the size of the
set of active columns is 40, one LOBPCG iterations implies doing at most 5
GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric}
and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES
iterations, using \pk{MatMult} and \pk{PCApply}. Results are gathered
in~\cref{tab:lobpcg}. In the first line, the numbers of outer LOBPCG iterations
are reported with the three different inner preconditioners mentioned above.
\pk{PCASM} is known to yield a preconditioned operator whose condition number
grows as the number of subdomains, here processes, increases. Thus, the number
of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the
inner solves are not converging to meaningful solutions in just 5 iterations.
The multilevel preconditioners perform similarly, since the difference in the
number of outer iterations is minimal (1). The second and third lines represent
the number of times \pk{PCApply} and \pk{PCMatApply} are called. Since
\pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration,
\pk{PCApply} is called on each single active column. \pk{PCASM} and
\pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}. In the fourth
line, the time spent setting up the preconditioners is given. \pk{PCASM} is
extremely cheap but not very robust numerically, while both multilevel
preconditioners have a similar setup time.  Eventually, in the last line, the
time spent in the full eigensolve is given.  Clearly, having both a \pk{KSP}
and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to
achieve reasonable performance with such an outer solver that heavily relies on
blocking.

@@

The script available at \url{https://github.com/prj-/jolivet2020petsc} may be
used to solve the same eigenproblem using \pk{EPSCISS} by providing the
additional command line argument \pk{-eps\_type ciss}. Blocking is also
available and tested in SLEPc test suite for \pk{EPSSUBSPACE} and
\pk{NEPCISS} since version 3.14.0.

@@

options is given in~\cref{fig:bratu-options}. Results can be reproduced using
the FreeFEM script \texttt{bratu.edp} available at
\url{https://github.com/prj-/jolivet2020petsc}.

@@

\section*{Acknowledgments}
The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
and B.~Smith for reviewing the successive merge requests in PETSc
repository. This work was granted access to the GENCI-sponsored HPC resources of:
\begin{itemize}
    \item TGCC@CEA under allocation A0070607519;
    \item IDRIS@CNRS under allocation AP010611780.
\end{itemize}
Jose E.~Roman was supported by the Spanish Agencia Estatal de Investigaci\'{o}n (AEI) under project SLEPc-DA (PID2019-107379RB-I00).
% We would like to acknowledge the assistance of volunteers in putting
% together this example manuscript and supplement.
